---
output: rmarkdown::github_document
---

#accel: R Functions to analyze accelerometer data

This package includes functions to read in and analyze accelerometer data generated by [Oregon Research Electronic AL100 Acceleration Loggers](https://sites.google.com/site/oregonresearchelectronics/home) to estimate tree phenology.

##Installation
```{r, eval = FALSE}
#Install from github
devtools::install_github("agougher/accel")
```

##Example data

The example data was generated by analyzing 250 acceleration files collected between March 1, 2016 and November 5, 2016, using the code below. The raw data was not included here because of its size, so the code is for illustrative purposes only. `readAndDom` reads in accelerometer files and calculates the dominant period of the three axes, as well as a weight.  
```{r, eval = FALSE}
#NOT RUN
#x is a list of ordered csv file locations
al3 <- lapply(x,readAndDom)
al3 <- do.call(rbind, al3)
al3["day"] <- 61:310
```

The example data includes the dominant period values for each of three dimensional axes (x, y, z), and weights reflective of the difference between the dominant and second-most dominant periods for each axis, for a single accelerometer.
```{r, message=FALSE}
#Load example data and make long form dataframe for functions below
require(accel) 
data(al3)
head(al3)

#Extract y and z axes, and bind them for the functions below
domY <- al3[,c("day","peakY","wpeakY")]
domZ <- al3[,c("day","peakZ","wpeakZ")]
colnames(domY) <- c("day","value","weights")
colnames(domZ) <- c("day","value","weights")
dat <- rbind(domY,domZ)

#Plot data, point size is proportional to the weight
plot(dat$day, dat$value, cex=dat$weights*2, xlab="Day of year",ylab="Dominant period")
```

##Removing outliers

This function, `rmOut`, identifies and removes points that have residuals beyond 1.5x the interquartile range of a loess curve. In the plot below, points in purple are kept after removing outliers.
```{r}
#The 'rmOut' function requires a dataframe with three columns named "day","value", and "weights"
dat2 <- rmOut(dat)
plot(dat$day, dat$value)
points(dat2$day, dat2$value, pch=20, col="purple", xlab="Day of year",ylab="Dominant period")
```

##Fitting phenology model described by [Elmore et al. 2012](http://onlinelibrary.wiley.com/doi/10.1111/j.1365-2486.2011.02521.x/full)

The model is a dual logistic curve, with an additional parameter that controls the slope of the line between the two logistic curves. In the plot below, black points are cleaned dominant period values, and red points are fitted values from the model. The two vertical lines are the spring and autumn inflection points of the model.

```{r}
#Remove NA's from data frame
dat2 <- na.omit(dat2)

#Set starting parameters
m <- c(20, 19, 145, 6, 275, 3, 1)

#Fit non-linear least squares model
mod = nls(value ~ a + (b- g*day)*((1/(1+exp((c-day)/d))) - (1/(1+exp((e-day)/f)))), 
             start = list(a = m[1],b = m[2],c = m[3], d = m[4], e = m[5], f = m[6], g = m[7] ), 
             algorithm="port",control = list(maxiter = 500), data=dat2, weights=dat2$weights^2)

#Plot data, and fitted values
#Units are in tenths of seconds, so divide by 10 to convert to seconds
plot(dat2$day, dat2$value/10, pch=20, xlab="Day of year",ylab="Dominant period")
points(dat2$day, fitted(mod)/10, pch=20, col="red", lwd=4)
abline(v=c(coef(mod)[3],coef(mod)[5]))
```

